---
title: "French given names per year per department"
author: "Edited by : **Oussama Oulkaid**. Template authors : **Lucas Mello Schnorr, Jean-Marc Vincent**"
date: "October, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[given names data set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004), we choose this dataset because it is sufficiently large, the analysis cannot be done by hand, the structure is simple.

We will use the _tidyverse_ for this analysis. The file **dpt2019.csv** contains the data. 

```{r, results=FALSE, message=FALSE}
# The environment
library(tidyverse)
library(ggplot2)
```


```{r, echo=FALSE, results=FALSE, message=FALSE}
version
```

<!-- Read in R with this code. Note that you might need to install the `readr` package with the appropriate command. -->

<!-- ## Download Raw Data from the website -->
```{r, echo=FALSE, eval=FALSE}
file = "dpt2020_txt.zip"
if(!file.exists(file)){
 download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2020_csv.zip",
 destfile=file)
 }
 unzip(file)
```

## Build the Dataframe from file
```{r, message=FALSE}
FirstNames <- read_delim("dpt2020.csv",delim =";")
```
```{r}
FirstNames
```

<!-- Translation in english of variables names:  -->
<!-- sexe -> gender  -->
<!-- preusuel (prénom usuel) -> Firstname -->
<!-- annais (année de naissance) -> Birth year -->
<!-- dpt (département) -> department (administrative area unit) -->
<!-- nombre -> number -->

<!-- All of these following questions may need a preliminary analysis of the data, feel free to present answers and justifications in your own order and structure your report as it should be for a scientific report. -->

<!-- 1. Choose a firstname and analyse its frequency along time. Compare several firstnames frequency -->
<!-- 2. Establish, by gender, the most given firstname by year.  -->
<!-- 3. Make a short synthesis -->
<!-- 4. Advanced (not mandatory) : is the firstname correlated with the localization (department) ? What could be a method to analyze such a correlation. -->

<!-- The report should be a pdf knitted from a  notebook (around  3 pages including figures), the notebook and the report should be delivered. -->

## Analysing first names frequencies
Let's choose a first name, and analyse its frequency.

```{r}
# FirstNames[sexe = c(1) , c('sexe','preusuel')]
# FirstNames[sexe = c('2'),]
# max(FirstNames$sexe)
albert <- subset(FirstNames[,c('preusuel', 'annais', 'nombre')], preusuel == 'ALBERT') %>% group_by(annais) %>% filter(annais!='XXXX')
#albert_annais <- albert_annais %>% summarise(nombre)
albert <- subset(albert[,c('annais', 'nombre')])
#albert[,c('nombre')] <- as.numeric(as.character( albert[,c('nombre')] ))
albert <- transform(albert, nombre = as.numeric(nombre), annais = as.numeric(annais))
albert <- setNames( aggregate(albert$nombre, by=list(annais=albert$annais), FUN=sum) , c('annais', 'sum_nombre') )
ggplot(data=albert, aes(x=annais, sum_nombre)) +
  geom_point() +
  geom_line(linetype = "dashed") +
  geom_rug(col="steelblue",alpha=0.1, size=1.5) +
  labs(title = "number of ALBERT names = f (year of birth)", x = "year of birth", y = "number of ALBERT names") +
  theme(plot.title = element_text(hjust = 0.5))
```
**Comment** : TODO

Now, let's compare several names' frequencies.

```{r}
# FirstNames[sexe = c(1) , c('sexe','preusuel')]
# FirstNames[sexe = c('2'),]
# max(FirstNames$sexe)
all_names <- subset(FirstNames[,c('preusuel', 'annais', 'nombre')]) %>% group_by(preusuel) %>% filter(annais!='XXXX')
sample_names <- subset( all_names[,c('preusuel', 'annais', 'nombre')], preusuel=='ALBERT' | preusuel=='CEDRIC' |preusuel=='_PRENOMS_RARES' | preusuel=='ARTHUR' )

sample_names <- transform(sample_names, nombre = as.numeric(nombre), annais = as.numeric(annais))
sample_names <- setNames( aggregate(sample_names$nombre, by=list(preusuel=sample_names$preusuel, annais=sample_names$annais), FUN=sum) , c('preusuel', 'annais', 'sum_nombre'))

ggplot(data=sample_names, aes(x=annais, sum_nombre, colour=preusuel)) +
  geom_point( size = 1 ) + #shape = "."
  geom_line(linetype = "dashed") +
  geom_rug(col="steelblue",alpha=0.1, size=1.5) +
  labs(title = "number of names = f (year of birth)", x = "year of birth", y = "number of names") +
  theme( plot.title = element_text(hjust = 0.5), legend.position = c(0.25, 0.75) )
```
---
title: "French given names per year per department"
author: "Edited by : **Oussama Oulkaid**. Template authors : **Lucas Mello Schnorr, Jean-Marc Vincent**"
date: "October, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[given names data set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004), we choose this dataset because it is sufficiently large, the analysis cannot be done by hand, the structure is simple.

We will use the _tidyverse_ for this analysis. The file **dpt2019.csv** contains the data. 

```{r, results=FALSE, message=FALSE}
# The environment
library(tidyverse)
library(ggplot2)
```


```{r, echo=FALSE, results=FALSE, message=FALSE}
version
```

<!-- Read in R with this code. Note that you might need to install the `readr` package with the appropriate command. -->

<!-- ## Download Raw Data from the website -->
```{r, echo=FALSE, eval=FALSE}
file = "dpt2020_txt.zip"
if(!file.exists(file)){
 download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2020_csv.zip",
 destfile=file)
 }
 unzip(file)
```

## Build the Dataframe from file
```{r, message=FALSE}
FirstNames <- read_delim("dpt2020.csv",delim =";")
```
```{r}
FirstNames
```

<!-- Translation in english of variables names:  -->
<!-- sexe -> gender  -->
<!-- preusuel (prénom usuel) -> Firstname -->
<!-- annais (année de naissance) -> Birth year -->
<!-- dpt (département) -> department (administrative area unit) -->
<!-- nombre -> number -->

<!-- All of these following questions may need a preliminary analysis of the data, feel free to present answers and justifications in your own order and structure your report as it should be for a scientific report. -->

<!-- 1. Choose a firstname and analyse its frequency along time. Compare several firstnames frequency -->
<!-- 2. Establish, by gender, the most given firstname by year.  -->
<!-- 3. Make a short synthesis -->
<!-- 4. Advanced (not mandatory) : is the firstname correlated with the localization (department) ? What could be a method to analyze such a correlation. -->

<!-- The report should be a pdf knitted from a  notebook (around  3 pages including figures), the notebook and the report should be delivered. -->

## Analysing first names frequencies
Let's choose a first name, and analyse its frequency.

```{r}
# FirstNames[sexe = c(1) , c('sexe','preusuel')]
# FirstNames[sexe = c('2'),]
# max(FirstNames$sexe)
albert <- subset(FirstNames[,c('preusuel', 'annais', 'nombre')], preusuel == 'ALBERT') %>% group_by(annais) %>% filter(annais!='XXXX')
#albert_annais <- albert_annais %>% summarise(nombre)
albert <- subset(albert[,c('annais', 'nombre')])
#albert[,c('nombre')] <- as.numeric(as.character( albert[,c('nombre')] ))
albert <- transform(albert, nombre = as.numeric(nombre), annais = as.numeric(annais))
albert <- setNames( aggregate(albert$nombre, by=list(annais=albert$annais), FUN=sum) , c('annais', 'sum_nombre') )
ggplot(data=albert, aes(x=annais, sum_nombre)) +
  geom_point() +
  geom_line(linetype = "dashed") +
  geom_rug(col="steelblue",alpha=0.1, size=1.5) +
  labs(title = "number of ALBERT names = f (year of birth)", x = "year of birth", y = "number of ALBERT names") +
  theme(plot.title = element_text(hjust = 0.5))
```
**Comment** : TODO

## Now, let's compute the most given firstname per year.

```{r}
# FirstNames[sexe = c(1) , c('sexe','preusuel')]
# FirstNames[sexe = c('2'),]
# max(FirstNames$sexe)
all_names <- subset(FirstNames[,c('preusuel', 'annais', 'nombre')]) %>% group_by(preusuel) %>% filter(annais!='XXXX')
#sample_names <- subset( all_names[,c('preusuel', 'annais', 'nombre')], preusuel=='ALBERT' | preusuel=='CEDRIC' |preusuel=='_PRENOMS_RARES' | preusuel=='ARTHUR' )
sample_names <- all_names[,c('preusuel', 'annais', 'nombre')]

sample_names <- transform(sample_names, nombre = as.numeric(nombre), annais = as.numeric(annais))
sample_names <- setNames( aggregate(sample_names$nombre, by=list(preusuel=sample_names$preusuel, annais=sample_names$annais), FUN=sum) , c('preusuel', 'annais', 'sum_nombre'))

#sample_names <- aggregate(sample_names$sum_nombre, by=list(annais=sample_names$annais), FUN=max)
#sample_names <- group[ group[, .I[sum_nombre == max(sum_nombre)], by=annais] ]
sample_names <- sample_names %>% group_by(annais) %>% top_n(1, sum_nombre)

ggplot(data=sample_names, aes(x=annais, sum_nombre, colour=preusuel)) +
  geom_point( size = 1 ) + #shape = "."
  geom_label( 
    data=sample_names, # Filter data first
    aes(label=preusuel)
  ) +
  geom_line(linetype = "dashed") +
  geom_rug(col="steelblue",alpha=0.1, size=1.5) +
  labs(title = "most given name = f (year of birth)", x = "year of birth", y = "most given name") +
  theme( plot.title = element_text(hjust = 0.5) )  #legend.position = c(0.74, 0.67) 
```















